### 2022-04-12 docs/MICRO 2022/details.tex
Add tiling section

\section{Optimizations}

\TODO{This section needs a better name!}
\input{tiling}