### 2022-04-15 docs/MICRO 2022/details.tex
Add prob tiling section

\section{Optimizations}

\TODO{This section needs a better name!}
\input{tiling}
\input{uniftiling}
\input{probtiling}