### 2022-04-15 docs/MICRO 2022/details.tex
Add interleaving, vectorization, parallelization

\section{Optimizations}

\TODO{This section needs a better name!}
\input{tiling}
\input{uniftiling}
\input{probtiling}
\input{unrollandjam}

\subsection{Vectorization}
Vectorization performed by Treebeard is enabled by the tiling transformations described in section \ref{sec:Tiling}. 
When the low level IR is translated to LLVM IR, Treebeard generates LLVM instructions that operate on the threholds and feature indices 
of nodes within a tile in a vector fashion. Therefore, thresholds and feature indices are loaded using vector loads and predicates are 
evaluated using vector comparisons. These vector LLVM IR instructions are then converted to vector instructions in the target ISA by 
the LLVM JIT.

\subsection{Parallelization}
Currently, Treebeard performs a naive parallelization of the inference computation. When parallelism is enabled, the 
loop over the input rows is parallelized using OpenMP. Treebeard rewrites 
the mid-level IR by tiling the loop over the input rows with a tile size equal to the number of cores. 
As a concrete example, consider the case where we intend to perform inference using a model with 4 trees 
on a batch of 64 rows. Further, assume that we wish to parallelize this computation across 8 cores. 
Treebeard then generates the following IR.
\begin{lstlisting}[style=c++]
  forest = ensemble(...)
  parallel.for i0 = 0 to 64 step 8 {
    for i1 = 0 to 8 step 1 {
      i = i0 + i1
      prediction = 0
      for t = 0 to 4 step 1 {
        tree = getTree(forest, t) 
        treePrediction = WalkTree(tree, rows[i])
        prediction = prediction + treePrediction
      }
      predictions[i] = prediction
    }
  }
\end{lstlisting}
Currently, as our focus is on single core performance, we do not 
do any parallelism related optimizations (such as loop tiling). We leave a more thorough exploration of parallelization 
to future work.