### 2022-04-17 docs/MICRO 2022/details.tex
Edits to tiling section

\section{Optimizations}

\TODO{This section needs a better name!}
\input{tiling}
\input{uniftiling}
\input{probtiling}
\input{unrollandjam}
\input{vectorization}

\subsection{Parallelization}
Currently, Treebeard performs a naive parallelization of the inference computation. When parallelism is enabled, the 
loop over the input rows is parallelized using OpenMP. Treebeard rewrites 
the mid-level IR by tiling the loop over the input rows with a tile size equal to the number of cores. 
As a concrete example, consider the case where we intend to perform inference using a model with 4 trees 
on a batch of 64 rows. Further, assume that we wish to parallelize this computation across 8 cores. 
Treebeard then generates the following IR.
\begin{lstlisting}[style=c++]
  forest = ensemble(...)
  parallel.for i0 = 0 to 64 step 8 {
    for i1 = 0 to 8 step 1 {
      i = i0 + i1
      prediction = 0
      for t = 0 to 4 step 1 {
        tree = getTree(forest, t) 
        treePrediction = WalkTree(tree, rows[i])
        prediction = prediction + treePrediction
      }
      predictions[i] = prediction
    }
  }
\end{lstlisting}
Currently, as our focus is on single core performance, we do not 
do any parallelism related optimizations (such as loop tiling). We leave a more thorough exploration of parallelization 
to future work.