### 2022-04-17 docs/MICRO 2022/probtiling.tex

\subsection{Probability Based Tiling}
\label{sec:ProbTiling}
\subsubsection{Motivation}
Uniform tiling optimizes tree walks assuming that all root to leaf paths in the tree are equally likely. 
However, we find that this is not always the case. We performed inference on the complete training set of each of our benchmarks and counted how many
times each leaf was reached. While for some benchmarks all leaves are roughly equally likely, for others, a small number of leaves are extremely likely 
while the rest aren't reached often. We call trees with a small number of extremely likely leaves \textbf{\emph{leaf biased}}.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/airline-ohe.stats.train.txt.png}
    \caption{Statistical profile for airline-ohe}
    \label{Fig:AirlineOHEStats}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/epsilon.stats.train.txt.png}
    \caption{Statistical profile for epsilon}
    \label{Fig:EpsilonStats}
\end{figure}

We say that an input row $r_i$ is \textbf{\emph{covered}} by a subset of leaves $L' \subseteq L$ of a tree $T$, if the leaf $l$ reached by 
walking $T$ for row $r_i$ is in $L'$. We show how different models (and even different trees within the same model) behave differently 
using models for two benchmarks, airline-ohe and epsilon. Consider the graphs shown in   
figures \ref{Fig:AirlineOHEStats} and \ref{Fig:EpsilonStats}. Each line in these graphs corresponds to a fixed fraction of the input (say $f$). 
A point on a line at coordinate $(x, y)$ means that a fraction $y$ of trees in the model could cover a fraction $f$ of all training inputs with a fraction $x$ of 
leaves. For example, the first point on the $f=0.9$ line in figure \ref{Fig:AirlineOHEStats} says that about 52\% of trees ($y$ value) need only 1\% of their
leaves ($x$ value) to cover 90\% of the training input. 
In general, Figure \ref{Fig:AirlineOHEStats} shows that very few leaves are needed to cover a very large fraction of inputs for the benchmark airline-ohe. 
This means that a small fraction of leaves are very likely. On the other hand, for the benchmark epsilon,
figure \ref{Fig:EpsilonStats} shows that a trees need a much larger fraction of their leaves to cover a significant fraction of the test input.
This means that most trees in the epsilon model are not leaf biased.
One other observation we make is that most models have some leaf biased trees while the rest of the trees have equally likely leaves.
\TODO{AP Maybe define a term for trees with roughly equally likely leaves?} We design the probability based tiling algorithm to take advantage of this property 
of decision tree ensembles.

\subsubsection{The Optimization Problem}

%We assume that we are given the probabilities of each leaf node of the decision tree (these can easily be computed using the training data). For every leaf $l \in L$, we %are given the probability $p_l$ that the leaf $l$ is reached.

We observe that the latency of one tree walk is proportional to the number of tiles that need to be evaluated to reach the leaf.
Therefore, the average inference latency can be modelled as the expected number of tiles that are evaluated to compute one tree prediction. More formally, the problem is to find a tiling $\mathcal{T}$ such that the following objective is minimized.
\[
    \min_{\mathcal{T} \in \mathcal{C}(T)}{\sum_{l \in L} p_l.depth_{\mathcal{T}}(l)}
\]
where the minimization is over all valid tilings $\mathcal{T}$ of the tree $T$, $depth_{\mathcal{T}}(l)$ is the depth of the leaf $l$ given tiling ${\mathcal{T}}$.

The constraints on the optimization are the constraints listed in Section \ref{sec:ValidTiling}. 
The above optimization problem can be solved optimally using dynamic programming. We leave this out in the interest of space. 
Instead, we use the simple greedy algorithm listed in algorithm \ref{Alg:GreedyTilingAlgo} to construct a valid tiling given the node probabilities.
The algorithm starts at the root and greedily keeps adding the most probable legal node to the current tile until the maximum tile size is reached.
Subsequently, the tiling procedure is recursively performed on all nodes that are destinations for edges going out of the constructed tile.

% \subsubsection{Dynamic Programming Formulation}

% For any node $v \in V$, we define
% \[
%     cost(v, \mathcal{T}) = \sum_{l \in L_v} p(l | v).depth_{\mathcal{T}}(l)
% \]
% where $\mathcal{T} \in \mathcal{C}(T_v)$.

% Then, the objective function, for the tree $T_v$, can be rewritten as 
% \[
%     opt\_cost(v) = \min_{\mathcal{T} \in \mathcal{C}(T_v)}{cost(v, \mathcal{T})}
% \]

% The objective function can then be rewritten in the following recursive form.
% \[
%     opt\_cost(v) = \min_{T_0 \in TileShapes(n_t, v)}{1 + \sum_{(n_1, n_2) \in out(T_0)} p(n_1 | v)p(n_2 | n1)opt\_cost(n_2)}
% \]
% where $TileShapes(n_t, v)$ is the set of all tile shapes of size $n_t$ with root $v$. A straight forward substitution argument shows why the solution to the subproblems (tiling all sub-trees) needs to to be optimal. The objective is now in a form that can solved using
% dynamic programming.

% \subsubsection{Greedy Algorithm}

% Intuitively, it seems like the following greedy algorithm also gives the optimal tiling. The algorithm starts at the root and greedily keeps adding the most probable node to the current tile until the maximum tile size is reached.
\begin{algorithm}
    \caption{Greedy Probability Based Tree Tiling}
    \label{Alg:GreedyTilingAlgo}
    \begin{algorithmic}
        \Procedure{TileTree}{$T = (V, E, r)$, $n_t$} 
            \If {$r \in L$}
                \State \textbf{return} $\{ r \}$
            \EndIf
            \State $Tile \leftarrow \{ r \}$
            \While{$|Tile| < n_t$}
                \State $e = (u,v) \in Out(Tile)$ st $p(v)$ is max and $v \notin L$
                \If{$e = \emptyset$}
                    \State \textbf{break}
                \EndIf
                \State $Tile = Tile \cup \{ v \}$
            \EndWhile
            \State $Tiles =  \{ Tile \}$
            \For{$(u,v) \in Out(Tile)$}
                \State $Tiles \leftarrow Tiles \cup TileTree(S_v, n_t)$
            \EndFor
            \State \textbf{return} $Tiles$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

% Talk about problems with increasing number of tile shapes and only performing such tiling on skewed trees
When we tried to apply algorithm \ref{Alg:GreedyTilingAlgo} on all trees in our benchmarks, we found
that even minor variations in probability caused the tiling algorithm to generate a large 
number of tile shapes. This in turn caused a loss in performance because the large size of the 
lookup table needed (section \ref{sec:LookupTable}) caused increased L1 cache misses. In order to 
alleviate this, we only perform probability based tiling on trees that are leaf biased. We decide that 
a tree is leaf biased if a small fraction of leaves, say $\alpha$, can cover a large fraction of training inputs, say $\beta$.
We only perform probability based tiling on skewed trees and fall back to uniform tiling otherwise. 
In our experiments we use $\alpha=0.05$ and $\beta=0.9$.