### 2022-04-18 docs/MICRO 2022/details.tex

%\section{Optimizations}
\section{high level IR Optimizations}
This section describes tiling and tree ordering, two optimizations performed at the highest level of abstraction. Recall that here the \op{predictForest} operator is abstractly represented by just a set of trees.

%\TODO{This section needs a better name!}
\input{tiling}
\input{uniftiling}
\input{probtiling}
\input{treeorder}
\section{Loop optimizations at mid level}
\input{unrollandjam}
\input{peelandunroll}
\input{parallel}
\section{Low level optimizations}

\input{vectorization}