### 2022-04-18 docs/MICRO 2022/related.tex

\section{Related Work}

\emph{Decision Tree Ensemble Compilers:}
Treelite\cite{Treelite} only supports generation of if-else style code and is not extensible.
Hummingbird\cite{Hummingbird} tries to compile decision tree models to tensor primitives so 
they can be integrated into tensor based frameworks like TensorFlow \cite{TensorFlow}.

\emph{Libraries:} Currently, the most popular systems for decision tree based models 
are libraries. XGBoost\cite{XGBoost} and LightGBM\cite{LightGBM} are the most  
popular gradient boosting libraries while scikit-learn\cite{Sklearn} is 
extremely popular for random forest models. These libraries implement both 
training and inference.

\emph{Other Systems and Techniques}: 
Ren et al \cite{PortableVM} build an intermediate language and VM to 
enable SIMD execution of decision tree inference. The SIMD execution itself is implemented 
by hand in the VM and the VM needs to be reimplemented for every supported target architecture.
Additionally, even though they perform layout optimizations, their system doesn't perform 
any model specific optimizations.

Tahoe, Vpred, CacheConscious1, QuickScorer, QuickScorer1, MilindTreeVectorization
Bliss?

Decision tree training parallelization references