### 2022-04-25 docs/MICRO 2022/conclusions.tex

\section{Conclusions}
This paper presents \Treebeard{}, a compiler that automatically generates  
efficient code for decision tree inference. \Treebeard{} gradually lowers  
inference code to LLVM IR through multiple intermediate abstractions.
It composes several novel optimizations to specialize inference code 
for each model on each supported hardware target. 
Experimental evaluation demonstrates that \Treebeard{} is 
significantly faster than state-of-the-art frameworks XGBoost and Treelite.

%% when executing 
% on both single core (by 2.6$\times$ and 4.7$\times$ respectively) and %% on 
% multiple cores (2.3$\times$ and 2.7$\times$ respectively). 
% The optimizations implemented in \Treebeard{} optimizations improved average latency over a batch of inputs by 2.2$\times$ over its
% own baseline version. %% lays a strong foundation to
%% automatically delivering high performance with high productivity through
%% compiler infrastructure for decision tree inference.
\Treebeard{} is a first step in automatically generating high performance
inference code for decision tree ensembles. It significantly simplifies retargeting 
high performance inference routines to new hardware targets. We believe that this methodology
can be extended to benefit a richer ensemble of ML models.