### 2022-04-25 docs/MICRO 2022/looporder.tex

\subsection{Loop Rewriting}
\label{Sec:LoopReordering}
\Treebeard{} supports a wide range of loop rewrites in the process of 
lowering from the high level IR to the mid level IR. \Treebeard{} 
supports fissing and permuting of the loop nest 
over the tree, input row pairs. Figure~\ref{Fig:Overview} shows the specific
case of permuting loops. Two versions of MIR generated for two different 
loop orders are shown -- ``one tree at a time'' that walks one tree for 
all input rows before moving to the next tree and ``one row at a time'' that
walks all trees for an input row before moving to the next row. The 
structure of the loop nest to be generated in MIR is decided at the HIR level and 
communicated to the lowering pass through attributes (as mention in 
section~\ref{Sec:Overview}).