### 2022-04-25 docs/MICRO 2022/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is the template for submission to MICRO 2022
% The cls file is modified from 'sig-alternate.cls'
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{sig-alternate}
\usepackage{mathptmx} % This is Times font
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage{tabularx,booktabs}
\usepackage{multirow}
\usepackage[normalem]{ulem}
\usepackage[hyphens]{url}
\usepackage[sort,nocompress]{cite}
\usepackage[final]{microtype}
\usepackage[keeplastbox]{flushend}
%\usepackage{subfig}
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{authblk}
% Always include hyperref last
\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}

\newcommand{\CommentOut}[1]{}
\newcommand\TODO[1]{\textcolor{red}{TODO #1}}
\newcommand\COMMENT[1]{\textcolor{green}{COMMENT #1}}
\newcommand{\Treebeard}{{\sc {Treebeard}}}
\newcommand{\op}[1]{{\texttt {#1}}}
\DeclareMathOperator{\Tree}{\ensuremath{\tau}}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{c++}{language=C++,
    morekeywords={ensemble, getTree, getRoot, isLeaf, isLeafTile, traverseTreeTile, getLeafValue, loadThresholds,
                  loadFeatureIndices, loadTileShape, getChildTile, parallel, InterleavedWalk, evaluateTilePredicates,
                  getChildTile, WalkDecisionTree, moveToChildTile},
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{blue},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\scriptsize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}

\lstset{style=c++}

% Ensure letter paper
\pdfpagewidth=8.5in
\pdfpageheight=11in

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\newcommand{\microsubmissionnumber}{654}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\fancypagestyle{firstpage}{
  \fancyhf{}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[C]{\vspace{10pt}\normalsize{MICRO 2022 Submission
      \textbf{\#\microsubmissionnumber} -- Confidential Draft -- Do NOT Distribute!!}\\\vspace{-25pt}}
  \fancyfoot[C]{\thepage}
}

\pagenumbering{arabic}

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\title{Treebeard : An Optimizing Compiler for Decision Tree Based ML Inference}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[1]{Ashwin Prasad}
\author[2]{Sampath Rajendra}
\author[2]{Kaushik Rajan}
\author[1]{R Govindarajan}
\author[1]{Uday Bondhugula}
\affil[1]{Department of Computer Science And Automation, Indian Institute of Science}
\affil[2]{Microsoft Research}

%\numberofauthors{5} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%

% \author{Ashwin Prasad}
% \affiliation{%
%  \institution{Indian Institute of Science}
%  \city{Bangalore}
%  \state{KA}
%  \country{India}}
% \email{ashwinprasad@iisc.ac.in}

% \author{
% % You can go ahead and credit any number of authors here,
% % e.g. one 'row of three' or two rows (consisting of one row of three
% % and a second row of one, two or three).
% %
% % The command \alignauthor (no curly braces needed) should
% % precede each author name, affiliation/snail-mail address and
% % e-mail address. Additionally, tag each line of
% % affiliation/address with \affaddr, and tag the
% % e-mail address with \email.
% %
% \alignauthor
% Ashwin Prasad\\
%       \affaddr{Indian Institute of Science, Bangalore}
%       \email{ashwinprasad@iisc.ac.in}
% \alignauthor
% Sampath Rajendra\\
%       \affaddr{Microsoft Research}
%        \email{srajendra@microsoft.com}
% \alignauthor
% Kaushik Rajan\\
%       \affaddr{Microsoft Research}
%        \email{krajan@microsoft.com}
% % \and
% \alignauthor
% R Govindarajan\\
%       \affaddr{Indian Institute of Science, Bangalore}
%       \email{govind@iisc.ac.in}
% \alignauthor
% Uday Bondhugula\\
%       \affaddr{Indian Institute of Science, Bangalore}
%       \email{udayb@iisc.ac.in}
% }

\renewcommand\Affilfont{\itshape\small}

\begin{document}
\maketitle
\thispagestyle{firstpage}
\pagestyle{plain}

%%%%%% -- PAPER CONTENT STARTS-- %%%%%%%%
\input{abstract}
\input{introduction}
\input{overview}
\input{opts}
\input{results}
\input{related}
\input{conclusions}

%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%

%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
\bibliographystyle{IEEEtranS}
\bibliography{refs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}