### 2024-03-29 docs/SOSP 2024/hir-mir-opts.tex
Add caching section

\section{HIR and MIR Optimizations}

\begin{itemize}
  \item The compiler performs a number of optimizations on the HIR and MIR to improve the performance of the generated code.
  \item Optimizations in HIR mainly transform the model itself. Tiling, tree reordering and padding are some of the 
  optimizations that are performed on the HIR. \TODO{Should we use the MICRO paper overview for this part?}
  \item Explain lowering to MIR using the schedule here.
  \item Optimizations on the MIR include tree walk unrolling, tree walk interleaving, loop tiling, parallelization etc.
\end{itemize}