### 2024-03-31 docs/SOSP 2024/caching.tex
Add sections for schedule space exploration and introduction

\section{Representations}
\TODO{Talk about array, sparse and reorg representations here.}

\section{Caching Implementation}
\label{sec:caching}

% \begin{itemize}
%   \item The compiler exposes caching of both trees and input rows in a unified manner.
%   \item This is independent of the final target on which the inference is to be run.
%   \item The caching is done at the granularity of a tree or a row.
%   \item Caching is encoded in the mid-level IR using the \op{cacheTrees} and \op{cacheRows} operations. 
%   These operations are generated when the HIR lowered to MIR and \op{Cache()} is specified on an 
%   index variable in the schedule. While the HIR is being lowered and a cached index variable is 
%   encountered, the compiler generates a \op{cacheTrees} or \op{cacheRows} operation depending on 
%   whether the index variable is a tree or a batch index variable.
%   \TODO{Should we talk about how the lowering from HIR to MIR is actually implemented as a tree walk?}
%   \item The lowering of these ops is done by the target-specific code generator.
%   \item For the \op{cacheRows} operation, \Treebeard{} uses pre-implemented lowerings for both
%   CPU and GPU. This is possible because the input is currently assumed to be a dense 
%   array format. 
%   \item For the \op{cacheTrees} operation, the lowering is representation-specific.
%   Each representation provides a lowering to the target-specific code generator
%   to lower the \op{cacheTrees} op when that representation is used.
%   \item The cache operations are lowered to reads into shared memory while compiling to GPU 
%   and to prefetches while compiling to CPU.
% \end{itemize}

\Treebeard{} provides mechanisms to cache both trees and input rows 
on both the CPU and GPU. As described in Section \ref{sec:schedule}, 
the user can specify that the working set of an iteration of an index
variable needs to be cached using the \op{Cache()} directive. This 
provides a unified way to specify caching of both trees and input rows.

\Treebeard{} implements caching at the granularity of a tree or a row.
Also, the semantics of caching depends on the target processor. For the CPU,
caching is implemented as prefetching, while for the GPU, caching is implemented
using shared memory.

\subsection{IR Representation of Caching}
Caching is encoded in the mid-level IR using the \op{cacheTrees} and \op{cacheRows} operations. 
These operations are generated when the HIR lowered to MIR and \op{Cache()} is specified on an 
index variable in the schedule. While the HIR is being lowered and a cached index variable is 
encountered, the compiler generates a \op{cacheTrees} or \op{cacheRows} operation depending on 
whether the index variable is a tree or a batch index variable. Additionally, as a part of the 
lowering process, \Treebeard{} determines the working set of the loop with caching enabled 
and generates a caching operation with the appropriate limits.

Each of the caching operations take parameters that specify the set of trees or rows that need 
to be cached. The caching operations are defined as follows. 
\begin{itemize}
  \item \textbf{\op{cacheTrees(forest, start, end)}:} This operation caches the trees in ensemble
  \op{forest} from \op{start} to \op{end}. The trees are cached in the order in which they are
  specified in the ensemble. 
  \item \textbf{\op{cacheRows(data, start, end)}:} This operation caches the rows in the input 
  array \op{data} from \op{start} to \op{end}. The rows are cached in the same order as in the
  input array.
\end{itemize}

\subsection{Lowering of Caching Operations}
When the MIR is lowered to LIR, the cache ops are lowered to target-specific code. Each of 
the two caching operations is lowered differently for the CPU and the GPU.

For the \op{cacheRows} operation, \Treebeard{} uses pre-implemented lowerings for both
CPU and GPU. This is possible because the input is currently assumed to be a dense
array format. Therefore, regardless of any other configuration choices (like what representation 
is used for the model itself), the lowering of the \op{cacheRows} operation is the same. 
For the CPU, the \op{cacheRows} operation is lowered to prefetches. For the GPU, a 
series of coalesced loads read the rows into shared memory.

For the \op{cacheTrees} operation, the lowering is representation-specific. Each representation
provides a lowering to the target-specific code generator to lower the \op{cacheTrees} op when
that representation is used. However, the \Treebeard{} infrastructure does provide helpers 
to generate caching code to cache contiguous regions of memory. These helpers are reused 
as required across different representations.

\section{Exploring the Schedule Space}
\label{sec:exploring}

\begin{itemize}
  \item The scheduling language described in Section \ref{sec:schedule} allows for numerous possible schedules for a given model. 
  Finding a schedule with good performance is a non-trivial task.
  \item We identify a reasonable template schedule for GPUs which encompasses all strategies published in prior work.
  \item Describe the template schedule.
  \item Even within the variants of this template schedule, there is a significant amount of variation in performance. Show the histogram. 
  \item Extensively exploring all possible parameter values for the template schedule is infeasible due to the large number of parameters.
  Cite some numbers on how long it takes to explore the schedule space.
  \item List the observations we make and the heuristic we design based on these.
\end{itemize}

The set of schedules that can be constructed using the scheduling language described in 
Section \ref{sec:schedule} is vast. Searching this schedule space to find a schedule that
provides good performance is a non-trivial task. To simplify this process, we identify a
reasonable template schedule for GPUs that encompasses all strategies published in prior work.
The template schedule exposes many parameters that can be tuned to find a 
high-performance schedule.
\begin{itemize}
  \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
  \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
  \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
  \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
  \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
  \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
  \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
  \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
\end{itemize}

While the template schedule simplifies optimization of generated inference code, it is 
important to note that the \Treebeard{} compiler itself does not place any 
restrictions on the schedule. The user is free to specify any schedule they wish.
The auto-scheduler that implements the template schedule is implemented as a 
module outside the core \Treebeard{} compiler. Users are also free to implement 
other auto-schedulers that generate schedules different from the template schedule.