### 2024-04-10 docs/SOSP 2024/abstract.tex
Edit abstract

\begin{abstract}

  The rapid proliferation of machine learning coupled with accelerating evolution of the hardware ecosystem has led to a surge in the demand for model inference on a variety of hardware.
  %While CPUs have been the mainstay for machine learning inference, the availability of GPUs holds promise to scale to bigger and more powerful models. 
  This paper is motivated by the problems encountered when targeting inference of decision tree based models, the most popular models on tabular data, to run at peak performance on 
  diverse CPU and GPU targets. We evaluated existing solutions and found that they do not provide portable performance across different hardware targets.
  %Decision tree based models are widely used in practice due to their robustness, interpretability, and ability to handle missing data.

  To address this we present the design of \Treebeard{}, a schedule guided compiler  
  for decision tree based models that searches over a large design space 
  to automatically generate high-performance inference routines.
  %We re-architect the open-source \TreebeardOLD{} infrastructure and significantly extend it to enable high-performance code generation across target processors. 
  \Treebeard{} has two core components. A scheduling language 
  that encapsulates the large optimization space for decision tree inference, 
  and techniques to efficiently explore this space.
  \TODO{Change large optimization space}.
  %We also design a set of heuristics that can find near-optimal solutions in quick time.
  Second, an optimizing multi-level compiler that can generate code based on the schedule. 
  For the latter, we re-architect the open-source \TreebeardOLD{} CPU compiler to support schedule guided compilation and
  add support for GPU code generation. GPU code generation required fundamental new optimization interfaces for 
  caching, parallel reduction, and support for multiple in-memory representations of trees.

  %The compiler builds on an existing compiler for CPUs, \TreebeardOLD{}, also add support for caching, parallel reduction and a plug-in mechanism to explore different in-memory representations of trees.

  We evaluate \Treebeard{} on over seven hundred diverse models and demonstrate that the best schedule varies drastically with model, batch size, and target hardware. 
  Our scheduling heuristic is able to quickly find near optimal schedules while searching over a small number (\~50) of schedules.
  In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
  On CPUs, \Treebeard{} achieves better scaling compared to \TreebeardOLD{}.
  For models where \TreebeardOLD{} was only able to achieve diminishing returns with an 
  increasing number of threads, \Treebeard{} is able to scale linearly with the number of threads.
  \TODO{(numbers for CPU performance?)}
\end{abstract}