### 2024-04-10 docs/SOSP 2024/intro.tex
Edit intro

\section{Introduction}
\label{sec:intro}
We are in the midst of a hardware revolution, a new golden age for computer architecture~\cite{GoldenAge}. The last 
decade has seen a shift in architectural paradigms, with the rise of GPUs and accelerators. This shift has been driven
by the necessity to innovate in the post Moore's law and Dennard's scaling era. This transformation has also played 
a significant role in the success of modern deep learning models, as they enable scaling model training and inference 
to a massive number of threads. Such scalability would be essential for all performance critical applications, including
other machine learning models that need to scale with increasing data sizes and model complexities. 
%need to accelerate machine learning workloads. The success of modern AI models is due to their ability to scale both training and inference to a massive number of threads. This has been made possible by the rise of GPUs and accelerators.

%Despite these recent advancements, one dominant family of models, namely decision forest models have not seen the same level of acceleration.
Decision forest models remain the mainstay for machine learning over tabular data~\cite{DLNotAllYouNeed,TreebasedOutperformDL}. 
Their robustness, interpretability, and ability to handle missing data make them a popular choice for a wide range of applications. 
\TODO{a few more lines, classification, regression, etc.}
%We observe that while scalable performance on neural network models has received significant attention, decision tree inference has not seen the same level of acceleration.
A recent survey~\cite{kaggle}\dots. 
The survey also finds that, the cost of inference is the most critical factor in the overall cost of deploying a machine learning model.
This is because, in production settings, each model is trained once and often used for inference millions of times. 
However, inference is run on a variety of hardware platforms, ranging from low to high-end CPUs and GPUs. 
This paper is motivated by the need to accelerate decision tree inference to achieve portable performance on a variety of hardware.
% In particular, we focus on a range of commodity CPUs and GPUs, a class of hardware that has seen widespread 
% adoption across client/edge devices used for inference.

Decision forest models are composed of a large collection of decision trees (100-1000), and inference involves 
traversing down each tree in the forest and aggregating the predictions. Inference is typically done in a batched
setting, where multiple inputs are processed simultaneously.
Despite the simplicity of the model and the availability of multiple sources of coarse grain parallelism (parallelism 
across inputs in a batch and parallelism across trees), existing systems do not consistently scale well across different
models even on the limited set of targets they support. 
This is because existing systems exploit a limited set of optimizations and often specialize the implementation 
to a specific hardware platform, and this limits their portability.

%\TODO{cover Rapids, Tahoe, XGBoost and Treebeard}.  
Evaluation on a diverse set of models highlights that the best implementation often requires a careful combination 
of many optimization strategies. For example, simultaneously traversing multiple trees with many threads, while managing the working set is 
critical to get scalable performance. This requires a combination of techniques like data layout optimizations, 
loop transformations, and memory access optimizations. Existing systems each use specific data layouts for the underlying
tree and use fixed loop structures. 
Parallelism, loop structure and model representation together determine the accesses patterns and the working set. A fixed configuration 
as used by existing systems does not perform consistently well across different models and hardware platforms.
XGBoost~\cite{XGBoost} uses a sparse representation for the model and a loop structure that processes one tree for a block of rows before moving to 
the next tree. RAPIDS FIL~\cite{FIL} uses a reorg representation and partitions trees across a fixed number of threads. Tahoe~\cite{Tahoe} uses a
variation of the reorg representation and has four predefined inference strategies from which it picks one based on an analytical model.
These systems are all library based and are hard to extend. 
\TreebeardOLD{}, the state-of-the-art model compiler for CPUs, supports two fixed loop structures and does not scale well with increasing number of threads.

This paper presents \Treebeard{}, a novel schedule guided compilation infrastructure for decision tree inference on multiple target hardware. \Treebeard{} is able to generate high-performance code for decision tree inference by exploring a large optimization space. The code generation is guided by a \emph{schedule} object that allows us to effectively represent this solution space abstractly. The schedule object is written in \Treebeard{} custom scheduling language that is expressive enough to represent a wide range of implementation strategies. We demonstrate that the language is sufficient to express the various optimizations proposed by prior work, and further generalizes them and incorporates several new optimizations. We also design and implement a heuristic that is able to quickly find high-performance schedules for the model being compiled.
\TODO{Performance evaluation summary}
\TODO{Oganization}
\subsection{Contributions}
\begin{itemize}
  \item We present the design for a multi-target decision tree compiler infrastructure and implement several optimizations within this framework.
   We are also the first to implement an optimizing compiler for decision tree inference on GPUs.
  \item We identify that an extensive optimization space exists for the problem of decision tree inference. We design a scheduling language that 
  allows us to effectively represent this solution space abstractly. This scheduling language is expressive enough to represent a wide range of  
  implementation strategies proposed by prior work.
  \item To the best of our knowledge, we perform the first extensive characterization of the optimization space for decision tree inference 
  on GPUs. Using some of the characteristics we identify, we design and implement a heuristic that is able to quickly find high-performance 
  schedules for the model being compiled. 
  \item We design and implement a general framework for expressing and optimizing reductions within MLIR. To the best of our knowledge, this is the first
  such framework.
  \item We evaluate our implementation by comparing it against RAPIDs and Tahoe, the state-of-the-art decision tree inference frameworks for GPU and 
  report significant speedups. We also show that our compiler can effectively target different GPUs, including both NVIDIA and AMD GPUs.
\end{itemize}