### 2024-04-11 docs/SOSP 2024/hir-mir-opts.tex
Edit schedule and hir-mir section

\section{HIR and MIR Optimizations}

% \begin{itemize}
%   \item The compiler performs a number of optimizations on the HIR and MIR to improve the performance of the generated code.
%   \item Optimizations in HIR mainly transform the model itself. Tiling, tree reordering and padding are some of the 
%   optimizations that are performed on the HIR. \TODO{Should we use the MICRO paper overview for this part?}
%   \item Explain lowering to MIR using the schedule here.
%   \item Optimizations on the MIR include tree walk unrolling, tree walk interleaving, loop tiling, parallelization etc.
% \end{itemize}

The \TreebeardOLD{} infrastructure was originally designed to target CPUs \cite{Treebeard}. 
However, it implements several optimizations on the HIR and MIR that 
can be leveraged across target processors and we find that some these 
are beneficial for GPUs as well. This reuse of optimizations is possible 
because the intermediate representations on which these optimizations are performed 
are abstract and are designed to be target-independent. We briefly review these 
optimizations below.

\subsection{Optimizations on High-Level IR}
We augment the existing \TreebeardOLD{} infrastructure with loop rewrites on 
the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).
We use these to implement the automatic scheduling described in Section \ref{sec:exploring}. 
Additionally, the \TreebeardOLD{} infrastructure implements HIR transformations to reorder and pad 
trees. It also implements tree tiling transformations on the HIR \cite{Treebeard}. We reuse the 
reordering and padding transformations on the HIR for GPUs. However, we found that 
tiling trees was not beneficial for GPUs. This is because the tiling transformation
introduces redundant computation inorder to vectorize computation on CPUs where 
all lanes need to follow the same control flow. However, on SIMT GPUs, we find that 
the benefits of tiling (coalescing memory accesses) do not outweigh the cost of
redundant computation. We leave an investigation of this for future work.

\subsection{Optimizations on Mid-Level IR}
The original \TreebeardOLD{} infrastructure implements optimizations like 
tree-walk unrolling, tree-walk interleaving, and parallelization
on the MIR. These optimizations are beneficial for GPUs as well.
and the design of \TreebeardOLD{} allows us to reuse the tree-walk 
unrolling and tree-walk interleaving optimizations on the MIR for GPUs.

While building \Treebeard{}, we found that one of the performance bottlenecks on the 
GPU was that warps spent significant time being stalled. Since GPUs 
implement scoreboarding~\cite{HennesseyPatterson}, we were able to alleviate this bottleneck by
interleaving tree walks. This significantly improved performance of generated 
code. We found it surprising that the use of ILP could benefit 
performance on the GPU.  
\TODO{Should we talk about how interleaving is implemented as a 
statemachine and therefore it can be used across representations 
and tile traversal techniques?}

\subsection{A Note on Low-level IR}
Significant changes to the original \TreebeardOLD{} design were required to get
LIR to correctly lower to GPU code. The most important of these was 
the change to how the compiler implements support for in-memory 
representations of models (Section \ref{sec:representations}).
With these design changes, we were able to reuse much of the CPU 
implementation while customizing some parts for GPUs (for example,
buffers need to be allocated differently for CPU and GPU, caching 
is implemented differently etc.).