### 2024-04-16 docs/SOSP 2024/schedule.tex

\section{Scheduling Language}
\label{sec:schedule}
As established in Section \ref{sec:motivation}, there are several 
different configurations and optimizations strategies for decision 
tree inference. The best ones vary significantly across
models, batch sizes and hardware platforms. Therefore, designing
any one hard-coded strategy is not feasible as this would 
make portable performance impossible. To address this, we 
design a scheduling language for \Treebeard{}. The scheduling 
language provides an abstract way to specify loop structure and 
other optimizations as an input to the compiler. The specified 
schedule controls the lowering of model inference to a set of 
loop nests. This configurability provided by the schedule allows us 
to build auto-schedulers and auto-tuners (Section \ref{sec:exploring}).
% Using a scheduling language will also significantly simplify 
% adding support for new hardware as this will likely require
% different locality optimizations and loop transformations.

% The goal of \Treebeard{}'s scheduling language is to declaratively
% express loop structures and the application of other optimizations 
% (tree walk unrolling, tree walk interleaving etc.).

% \subsection{Language Definition}
The core construct of \Treebeard{}'s scheduling language is an 
\textbf{\emph{index variable}} which abstractly represents a loop. 
The language then provides directives to manipulate these index 
variables. There are two special index variables -- \op{batch} and
\op{tree} that are used to represent the batch and tree loops and all 
other index variables are derived from these. A schedule derives 
new index variables from these root index variables by applying
directives.

\Treebeard{}'s scheduling language has three classes of directives. The first is a set 
of loop modifiers that are used to specify the structure of the loop nest to
walk the iteration space (Table \ref{Tab:LoopModifiers}). The second is a set of 
directives that enable optimizations on a loop (Table \ref{Tab:Optimizations}). 
Finally, we have a class of attributes that enable reduction specific optimizations
(Table \ref{Tab:ReductionOpts}).

The set of loops (index variables) are internally represented as nodes in a tree where the children
of a node represent immediately contained loops. Each schedule 
primitive modifies this tree in some way. The compiler tracks 
the lineage of each of the loops. This allows the compiler to 
automatically infer the ranges for all loops.

% \subsubsection{Loop Modifiers}
% The clauses modify these index variables or 
% index variables derived from these (through the application of clauses).
% \begin{itemize}
%   \item \textbf{tile}: Tile the passed index variable using a fixed tile size.
%   \item \textbf{split}: Split the range of the passed index variable into two parts. The range of the first part is specified
%   by an argument.
%   % \item \textbf{unroll}: Unroll an index completely
%   \item \textbf{reorder}: Reorder the specified indices. The specified indices must be successive indices in the current loop nest.
%   \item \textbf{specialize}: Generate separate code for each iteration of the specified index variable. This is useful 
%   while parallelizing across trees and these trees have different depths.
%   \item \textbf{gpuDimension}: Maps the specified index variable to represent a dimension in either the GPU kernel grid or thread block.
% \end{itemize}

\begin{table}[htb]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabularx}{\linewidth}{c | l | l}
   \toprule
   \textbf{Directive} & \textbf{Inputs} & \textbf{Description} \\
   \midrule
   \multirow{4}{*}{\texttt{tile}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Tile the loop corresponding to \textbf{indexVar}
   with the specified tile size. Resulting loops will be represented by \textbf{outer} and \textbf{inner}.}} \\
                                               &  \textbf{outer} & \\
                                               &  \textbf{inner} &  \\
                                               &  \textbf{tileSize} &  \\
   \midrule
   \multirow{6}{*}{\texttt{split}} & \textbf{indexVar}  & \multirow{6}{*}{\parbox{0.55\linewidth}{Fiss the loop represented by \textbf{indexVar}
   at iteration \textbf{splitIter}. Resulting loops will be represented by \textbf{first} and \textbf{second}. Returns a maps from nested 
   index variables to new ones created by splitting.}} \\
                                               &  \textbf{first} & \\
                                               &  \textbf{second} &  \\
                                               &  \textbf{splitIter} &  \\
                                               & & \\
                                               & & \\

   \midrule
   \multirow{4}{*}{\texttt{reorder}} & \textbf{indices[]}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Permute loops corresponding to the specified index variables.
   The loops must be perfectly nested in the current loop structure.}} \\
                                               &   & \\
                                               &   &  \\
                                               &   &  \\
  %  \midrule
  %  \multirow{4}{*}{\texttt{specialize}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Generate specialized code for each iteration of the loop. 
  %  Useful when different iterations of a loop need to execute different code.}} \\
  %                                              &   & \\
  %                                              &   &  \\
  %                                              &   &  \\
   \midrule
   \multirow{3}{*}{\texttt{gpuDimension}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Map the passed index variable to a dimension of either the grid
   or thread block.}} \\
                                               & \textbf{gpuDim}  & \\
                                               &   &  \\

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:LoopModifiers} List of all the loop modifiers in \Treebeard{}'s scheduling language. We use \emph{index variable}
  and \emph{loop} interchangeably in descriptions for clarity of exposition.}
\end{table}

% The following are examples of how the loop modifiers can be used.
% \begin{itemize}
%   \item The loop order used by XGBoost\cite{XGBoost} is (tree, batch) -- walk one tree for all inputs in the batch before 
%   moving to the next tree. The corresponding schedule would be
% \begin{lstlisting}[style=c++]
%   reorder(tree, batch)
% \end{lstlisting}

%   \item The below schedule computes 2 trees at a time over the whole batch.
% \begin{lstlisting}[style=c++]
%   tile(tree, t0, t1, 2)
%   reorder(t0, batch, t1)
% \end{lstlisting}

%   \item If we additionally only want to compute over 4 input rows (rather than the whole batch) for
%   every 2 tree, and then move onto the next 2 trees for the same set of inputs, then the schedule is as follows. 
% \begin{lstlisting}[style=c++]
%   tile(batch, b0, b1, 4) 
%   tile(tree, t0, t1, 2)
%   reorder(b0, t0, b1, t1) 
% \end{lstlisting}

% \end{itemize}

% \subsubsection{Optimizations}
% The following clauses provide ways to optimize the inference routine being generated.
% \begin{itemize}
%   \item \textbf{cache}: Cache the working set of one iteration of the specified loop corresponding
%   to this index. This can be specified on either batch or tree loops. Specifying it 
%   on a batch loop leads to all rows accessed in a single iteration of the loop 
%   being cached. Similarly, specifying it on a tree loop leads to all trees accessed in 
%   one iteration of that loop being cached.
%   \item \textbf{parallel}: Execute the loop corresponding to this index in parallel.
%   \item \textbf{interleave}: Interleave the execution of the tree walks within the current index (must be applied on an inner most index).
%   \item \textbf{unrollWalk}: Unroll tree walks at the current index. 
%   \item \textbf{peelWalk}: Peel the first n steps of the specified tree walk and don't check for leaves for that number of steps.
% \end{itemize}

\begin{table}[htb]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabularx}{\linewidth}{c | l | l}
   \toprule
   \textbf{Directive} & \textbf{Inputs} & \textbf{Description} \\
   \midrule
   \multirow{4}{*}{\texttt{cache}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Cache the working set of one iteration of the specified loop. 
   Cache rows for a batch loop and trees for a tree loop.}} \\
                                               &  &  \\
                                               &  &  \\
                                               &  &  \\
   \midrule
   \multirow{2}{*}{\texttt{parallel}} & \textbf{indexVar}  & \multirow{2}{*}{\parbox{0.55\linewidth}{Execute the iterations of the specified 
   loop in parallel.}} \\
                                               &  & \\

   \midrule                                               
   \multirow{3}{*}{\texttt{interleave}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Interleave tree walks within the specified 
   loop (must be innermost loop).}} \\
                                               &  & \\
                                               &  & \\

   \midrule                                               
   \multirow{3}{*}{\texttt{unrollWalk}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Unroll tree walks at the specified loop 
   for \textbf{unrollDepth} hops. Loop must be an innermost loop.}} \\
                                               & \textbf{unrollDepth} & \\
                                               &  & \\

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:Optimizations} List of optimization directives in \Treebeard{}'s scheduling language. 
  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
\end{table}

% \subsubsection{Reduction Optimization}
% \begin{itemize}
%   \item \textbf{atomicReduce}: Use atomic memory operations to accumulate values across 
%   parallel iterations of the specified loop. 
%   \item \textbf{sharedReduce}: Only applies to GPU compilation. Specifies that intermediate
%   results are to be stored in shared memory.
%   \item \textbf{vectorReduce}: Use vector instructions with the specified vector width 
%   to reduce intermediate values across parallel iterations of the specified loop.
% \end{itemize}

\begin{table}[htb]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabularx}{\linewidth}{c | l | l}
   \toprule
   \textbf{Directive} & \textbf{Inputs} & \textbf{Description} \\
   \midrule
   \multirow{3}{*}{\texttt{atomicReduce}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Use atomic memory operations to accumulate values across 
   parallel iterations of the specified loop.}} \\
                                               &  &  \\
                                               &  &  \\
   \midrule
   \multirow{3}{*}{\texttt{sharedReduce}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Specifies that intermediate
   results are to be stored in shared memory (GPU only).}} \\
                                               &  & \\
                                               &  & \\

   \midrule                                               
   \multirow{4}{*}{\texttt{vectorReduce}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Use vector instructions with the specified vector width 
   to reduce intermediate values across parallel iterations of the specified loop.}} \\
                                               & \textbf{width} & \\
                                               &  & \\
                                               &  & \\

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:ReductionOpts} List of reduction optimization directives in \Treebeard{}'s scheduling language. 
  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
\end{table}

%\TODO{Add examples for RAPIDS, Tahoe (Maybe show some strategies can be encoded?)}

\Treebeard{}'s scheduling language is expressive enough to represent a wide range of
strategies used in existing systems. We show examples of how it can be used to 
represent XGBoost and Tahoe's strategies. Before presenting the examples, we note that \Treebeard{}'s 
default loop order is [\op{batch}, \op{tree}], i.e, for each row in the input
batch, go over all trees.

% \subsection{The XBoost Schedule}
XGBoost\cite{XGBoost} implements inference on the CPU by going 
over a fixed number of rows (64 in the previous version)
for every tree and then moving to the next tree. When all trees 
have been walked for this set of rows, the next set of rows is 
taken up. Different sets of rows are processed in parallel.
The following schedule expresses XGBoost's strategy.
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, CHUNK_SIZE)
  reorder(b0, tree, b1)
  parallel(b0)
\end{lstlisting}

Tahoe\cite{Tahoe} has four strategies for inference on the GPU that it picks from for a given model. 
We show how two of these strategies can be encoded using \Treebeard{}'s scheduling language.
%The rest can be encoded similarly. 
\begin{itemize}
  \item In the \emph{direct method}, a single GPU thread walks all trees
  for a given input row. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, ROWS_PER_TB)
  reorder(b0, b1, tree)
  gpuDimension(b0, grid.x)
  gpuDimension(b1, block.x)
\end{lstlisting}
  Here, \op{ROWS\_PER\_TB} is the number of rows that are processed by a single thread block.
  \item In the \emph{shared data} strategy, a thread block walks all the trees 
  for a given row in parallel. 
  % If threads walk multiple trees, each thread accumulates partial results. 
  Then, a thread block wide reduction is performed to compute 
  the prediction. The schedule for this strategy is as follows.
\begin{lstlisting}[style=c++]
  reorder(batch, tree)
  gpuDimension(batch, grid.x)
  gpuDimension(tree, block.x)
  cache(batch)
\end{lstlisting}
%   \item \textbf{Shared Forest}: In this strategy, the whole model is loaded into 
%   shared memory and subsequently, a single thread walks all trees for a particular
%   row. The schedule for this strategy is as follows.
% \begin{lstlisting}[style=c++]
%   tile(batch, b0, b1, ROWS_PER_TB)
%   tile(tree, t0, t1, N_TREES)
%   reorder(b0, b1, t0, t1)
%   cache(t0)
%   gpuDimension(b0, grid.x)
%   gpuDimension(b1, block.x)
% \end{lstlisting}
%   Here, we create a placeholder single iteration loop \op{t0} so that we can 
%   specify that all trees are to be cached.
%   \item \textbf{Shared Partial Forest}: In case the model is too large to fit into
%   shared memory, the model is split into chunks and each chunk is loaded into shared
%   memory. Again, as in the previous strategy, one thread walks all trees assigned to
%   a thread block for a row. The schedule for this strategy is as follows.
% \begin{lstlisting}[style=c++]
%   tile(batch, b0, b1, ROWS_PER_TB)

%   tile(tree, t0, t0Inner, TREES_PER_TB)
%   tile(t0Inner, t1, t2, TREES_PER_TB)
%   cache(t1)
%   reorder(b0, t0, b1, t1, t2)

%   gpuDimension(b0, grid.x)
%   gpuDimension(t0, grid.y)
%   gpuDimension(b1, block.x)
% \end{lstlisting}
\end{itemize}

% There are some simplifying assumptions and limitations in the current design
% of the scheduling language. Mainly, tree traversals are considered atomic and
% accumulation of tree predictions is done immediately (as opposed to, 
% for example, collecting all predictions and performing a reduction later).
% However, we find that these are not significant limitations in practice as
% the current design is able to express most strategies of interest.